---
title: "02-jan11"
format: gfm
editor: visual
---

```{r setup}
library(tidyverse)
library(ggplot2)
options(mc.cores = parallel::detectCores(logical=FALSE))
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file())
setwd(rprojroot::find_rstudio_root_file())
```

```{r}

vec <- 1:10
A <- 1:10
B <- seq(from = 1, to = 2, length = 10)
C <- rep(1,10)
D <- cbind(A,B,C)

D <- LETTERS[1:10]

E <- cbind(A,B,C, D) 

# change everything in characters bc vector/matrix changes everything into same class

E <- list(A,B,C,D, F = 1:25)
# list has 2 brackets ]] 
# list allows dif. class

# you can either access with brackets [[]] for the columns, 
# or the $ is you assigned a name 

# what is a data.frame ?
# mixed of list and matrix = allows dif. classes but all the elements are the same length 

G <- data.frame(A=A, B=B, C=C)

# look at the object mode or class

#to select 3 cols only, with all the rows: 
m <- as.matrix( G[,1:3])
m <- as.matrix( G[,c("A", "B", "C")]) # i personnally prefer this bc it calls the actual name, not the nbrs, so less change of a mistake  

#array s

K <- array(data = NA, dim = c(2,3,4))

#reshape package and melt function
# check nest data 

# you can use function scan  to read a series of numbers
# - helpful to reading crazy data file from old pdf ... like ebmd

#LOOOOOOPPPPSSS !!!!!!!!
#creating a vector 
B <- rep(NA,9)
for( i in 2:10){
  B[i] <- A[i-1] +3
}

```

## Creating model

expGrowth

written by Brett

Jan 11,2023

This code estimates exponential growth for lakes in 7 regions in BC, it returns tgh number in each region where it declines occur.

*pseudo code* :

( what is pseudocode : step by step description of the coding )

1\. set the working directory

2\. copy count data into the working directory

3\. input the data using read.table

4\. set DecliningPopCount to Zero

5\. Loop over lakes

5a. extract NT and N0 observations

5b. compute growth rate 5c. increment DecliningPopCount if negative growth rate

6\. report decliningPopCount using print() or cat()

```{r creating model}
#this is incase you dont use github and project folders
#this library allows you to write to path
install.packages("here")
library(here)

#path <- here() 

#save data 
data <- read.table(file = "data/lecture2/Counts1.dat", header = TRUE, sep = "")

#decliningPopCount is a counter for the nbrs of declining poopulation

decliningpopcount <- 0
#determine number of lakes
nlake <- nrow(data)
#nbrs of years between observation 
nyr <- 10

#loop over lakes
for (i in 1:nlake){
  NT <- data[i,'NT']
  N0 <- data[i,'N0']
  g <- log(NT / N0)/ nyr
  if( g < 0) decliningpopcount <- decliningpopcount + 1
}

cat("the number of lakes with declining abundance is", decliningpopcount)
```
