---
title: "06-feb06"
format: gfm
editor: visual
---

##Weeek 6 :

```{r setup}
library(tidyverse)
library(cowplot); theme_set(theme_cowplot())
library(patchwork)


options(mc.cores = parallel::detectCores(logical=FALSE))
knitr::opts_knit$set(root.dir = rprojroot::find_rstudio_root_file()) 
setwd(rprojroot::find_rstudio_root_file())

```

##logistic growth this is an equalibirum model - so there is a looot of assumptions r = maxium rate of growth K = carrying capacity

equation

$$
\frac{d_r}{d_t} = rN(1-\frac{N}{K} )
$$

$$
N_t+1 = N_t +rN_t(1-\frac{N}{K} )
$$

growth rate slow down = difference between N and N+1 diminish as the population growth

at the top of the logistical grwth , that s where the (NT+1 - Nt) is the biggest!higgest recruitment! how to calculate !! slope of r (growth r) is positive until it is negative so we need to find when the slope of growth is 0 So we take the derivative 
$$
rN - \frac{r}{k}N^2 = 0
$$
SOOOOOOO it ends up being : 
$$
N_msy = K/2
$$

at equlibrium, N t+1 = N t
$$
U_{msy} = \frac{r}{2}
$$

How high is the curve ( aka what is the $\delta$N)


```{r}



```
